Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence and Forensic Workflow

### 2009-03-13 paper/paper.tex
More work on paper.

@@

\section{Streams}
The Stream object provides random access to an abstract representation
of a body of data. Our implementation allows the segments in a stream
to be operated on as if they were a single file by supporting the
traditional POSIX-like functionality of
\texttt{open()}, \texttt{seek()}, \texttt{write()}, and \texttt{read()}.

@@

\subsection{The Image Stream}
\label{image_stream}
The AFF4 \emph{Image Stream} stores a single read-only forensic data
set. For example, this stream might contains a hard disk image, a
memory image or a network capture (in PCAP format).

Storage for the data is done by using multiple data segments stored on
various volumes. Data segment URNs are derived by appending an 8
digit, zero padded decimal integer representation of an incrementing
id to the stream URN (e.g. ``urn:aff2:83a3d6db-85d5/00000032'').

Image streams must specify the
\texttt{page\_size} attribute, as the number of image bytes each chunk
contains. Pages may be compressed at the page level, or may be broken
up into multiple chunks (typically 8KB), and individually
compressed. In this case the beginning of the page segment will
contain an offset table indicating the location of each compression
chunk. This is similar to the technique used by the Expert Witness
file format used by EnCase\cite{encase-3.0} and implemented by the
open source libewf\cite{libewf} package.

Pages can optionally be individually hashed, allowing for detection
of modifications. The attribute \texttt{hash\_type} specifies the type of
hash used (implementations will typically support md5,sha1 and sha256, but can
define others). The hashes are stored in any number of segments within
the stream, of arbitrary name. The hash segment names is provided by
using the \texttt{hash\_name} stream attribute.

Image streams must have at several attributes. First, they must have
\texttt{type=image} to indicate that they are image streams. Streams
must have a \texttt{size} attribute (specified in 8-bit bytes) to
indicate the last addressable byte in the stream. Streams can have
holes in them; in this case, attempts to read data from within the
holes can generate an error or return NULLs.

Hashes are calculated in an identical way to that defined by AFF1's
crypto layer\cite{garfinkel:affcrypto}---that is, the full name of the
segment (including stream component) is followed by a single NULL
byte, followed by the (uncompressed) content of the chunk. The
inclusion of the segment name ensure that even identical chunks in
content will have a different hash and prevents reordering attacks.

Each hash segment is a list of a chunk id encoded in 32 bit
little endian integer format followed by the hash (the length of which
depends on the hash itself). This approach is illustrated in Figure
\ref{hash_index}.

Hash segments can contain any number of chunk hashes in them. It is
recommended that a hash segment be created when a volume is closed to
cover all the chunks within that volume. This allows for hashes to be
verified even if some volumes are missing. The AFFLIB implementation
may update a hash of a chunk previously set by an earlier hash
segment---this would be required when a chunk is updated. The hash
segment can then be signed to assure the integrity of the updated segment.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
  \caption{The structure of a hashing segment. The attributes {\em
  hash} and {\em hash\_type} declare the hash segment. The segment
  contains a list of integer encoded chunk id followed by an unencoded
  hash for each chunk.}
  \label{hash_index}
  \end{center}
\end{figure}

This design differs from AFF1, which creates a separate segment for
each hash. AFF4 uses headers that are more verbose than the AFF
segment headers. Therefore, creating a new segment for each hash value
would result in too much overhead: Packing the hashes into a few
segments it is possible to reduce this overhead to a minimum.

\subsection{The Map Stream}
\label{map_stream}
Linear transformations of data are commonplace in forensic
analysis. For example, a file is often simply a collection of bytes
drawn from an image, while a TCP/IP stream is simply a collection of
payloads from selected network packets. 
Sometimes the same data may be
viewed in a number of ways---for example a Virtual Address Space is a
mapping of the Physical Address Space through a page table
transformation.  Zero Storage Carving \cite{Meijer2006} is a way of
specifying carved files in terms of a sequence of blocks taken from
the image; Cohen extended this concept to an arbitrary
mapping function\cite{Cohen2007} .

AFF4 defines a common way in
which the underlying storage system can apply transformations to
existing data to produce transformed data. This transformation is
performed using a \emph{map stream}.

The Map stream defines a mapping function, as a piecewise continuous
linear transform of one or more {\em target} streams to produce a new
{\em stream}. Target streams are specified by using ``target''
attribute, which may be used many times to provide a list of targets.

The mapping function file consists of a series of lines, each
containing three 64 bit integers separated by comma, and encoded using
decimal notation. The integers represent stream offset, target offset
and target number. Target numbers are an index to the list of targets
given in the attribute section (list index starts at zero for the
first entry).

@@

For example, consider the following map:
\begin{verbatim*}
0,0,0
4096,10000,0
8192,5000,0
\end{verbatim*}

To read this stream we satisfy read requests of offsets between 0 and
4095 in the stream from offset 0 in target 0. Requests for bytes
between 4096 and 8191 are fetched from target 0 at offset
10000. Finally bytes after 8192 (until the specified size of the
stream) are fetched from offset 5000 in target 0.

@@

Where $mod$ is the modulus function and $floor$ signifies integer
division. For example consider Figure~\ref{map}, which corresponds to a 3
disk RAID-5 array.

\begin{lstlisting}[
	float=tb, caption=A Map stream that corresponds to a 3 disk
	RAID-5 array. The targets are URNs for the respective
        disks. Note that map coordinates are given in multiples of 
	block size.,
	label=ref, language=]
aff4:block_size=64k 
aff4:stream_period=6 
aff4:target_period=3

@@

\subsection{The Overlay Stream}
The Overlay Stream is a variant of the {\em Image} stream that
provides a layer of indirection between requests made for specific
bytes (or chunks or bevys) and the actual location where the requested
information is stored. The main utility of the Overlay stream is in
providing transparent access to legacy forensic formats, such as EWF.

To access these legacy file formats we only need to have an index of
each chunk offset, and then directly use the overlayed files.  The {\em
ewf2aff} tool provides such an overlay feature for EWF volumes by
employing libewf to scan the internal data structures.

Overlay stream require that a \texttt{chunk\_size} attribute be specified. In
addition the {\em compression\_type} attribute may be specified. The
attribute should contain an integer corresponding to a suitable Zip
compression scheme (e.g. DEFLATE). If not specified it is assumed
that no compression is used.

The ``target'' stream attribute may be specified multiple times and
refer to external files (See section \ref{target_reference}). There
can be a number of overlay segments named ``stream name/overlay.00''
which are all merged to a single overlay. If more than one overlay is
present, the attribute {\em overlays} MUST be specify how many.

The overlay segment contains a series of lines with comma delimited,
decimal encoded, 64 bit integers representing chunk number, target
offset, target length and target number. Target number is an index
into the targets defined in the attributes (count starts with zero).
Chunks are then fetched from the specified target and decompressed if
necessary.

It is possible to refer to the current file by providing a target of
``file://.''. This can be useful for modifying an existing EWF file to
be a AFF4 file. Since Zip files are normally read from the end of the
file, but EWF are both read from the front of the file, it may be
possible to append a AFF4 Central Directory to the end of an EWF file
without interfering with the overlayed file. In this case the
converted file can still be used as an EWF file without change, but
could also be used as an AFF4 file.

In image streams some chunks may compress very well. In that case, the
overheads introduced by the Zip File Header for each chunk in the
Image stream may become unacceptably high. In that case its possible
to use an {\em Overlay stream} to coalesce chunks into larger
segments. This approach is illustrated in Figure \ref{overlay}.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{overlay.eps}}
  \caption{The use of an Overlay Stream to overlay an image with
coalesced chunks.}
  \label{overlay}
  \end{center}
\end{figure}

In the above figure a segment is written to the AFF4 volume which
contains a number of chunks back to back. An overlay stream then uses
direct references to each compressed chunk within the coalesced
segment. The target is then specified as the segment name. Although
this technique is valid, it is generally discouraged since it breaks
the appealing simplicity of the standard Image Stream.  It is no
longer possible to simply unzip all the chunks and concatenate them
together to recover the original image.

\subsection{Encrypted Streams}
Encryption is an important property in an evidence file format. In
particular, multiple streams may be present in the file set, and often
different access levels are desired. For example, for evidence set
containing both network captures and disk images it may be desirable
to limit access to streams based on legal authorizations.

@@

AFF4 therefore introduces a new encryption scheme, the Encrypted
Stream.  The Encrypted Stream is a stream that encapsulates other
streams on a bevy-by-bevy basis. Each bevy transparently encrypted upon
writing and decrypted on reading. The choice of algorithm and mode is declared in the
Stream's \texttt{scheme} attribute.

The Encrypted Stream may contain any data at all. It is useful
however, to store an actual AFF4 volume within the Encrypted
stream. This provides block level encryption for the contained AFF4
volume. It is recommended that Encrypted Streams containing a AFF
volume set an attribute (\texttt{content-type=application/x-aff-volume})
to allow implementations to automatically recognize volumes. This
approach is illustrated in Figure~\ref{crypted_fif}.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{crypted.eps}}
  \caption{Embedding an encrypted AFF4 volume within an Encrypted Stream.}
  \label{crypted_fif}
  \end{center}
\end{figure}

@@

\subsection{Encryption Schemes}
We specify a number of encryption schemes, but extra schemes may be
defined by the implementation.

@@

When creating a new volume, a master key is generated using the first
128 bits of a SHA1 hash of the Pre-Shared-Key appended to a 64 bit
random salt:
\begin{eqnarray}
Key_{Master} = \left | SHA1(salt + PSK) \right | _{128}
\end{eqnarray}

@@

The block is then encrypted or decrypted using AES with this IV.

@@

\item Our implementation supports simultaneous access by multiple
  readers and writers. Our system allows concurrency by locking the
  file and interleaving write operations. Managing simultaneous
  writers is simplified through our update strategy which only writes
  new segments to the file, never modifying old segments. 
\item AFF4 ``properties'' have specific semantics above and beyond
  normal segments. For example, AFF4 calls for all of the property
  segments associated with a file to be merged together. Furthermore,
  when updating the properties in a multi-volume Set, it is useful to
  write the same properties to each volume to ensure at least some of
the stream can be used even if some volumes are lost. Such
functionality is not present in standard ZIP64 implementations.
\end{itemize}

>>>>

Each URN within the AFF4 universe must have an ``\texttt{type}''
attribute to denote the type of the Object.

@@

\section{Streams}
The Stream object provides random access to an abstract representation
of a body of data. Our implementation allows the segments in a stream
to be operated on as if they were a single file by supporting the
traditional POSIX-like functionality of
\texttt{open()}, \texttt{seek()}, \texttt{write()}, and
\texttt{read()}. All streams must have the ``\texttt{size}'' attribute
to denote the last bytes addressible within the stream. This is
required in order to support the POSIX whence attribute which may
require seeking from the end of the stream.

@@

\subsection{The Image Stream}
\label{image_stream}
The AFF4 \emph{Image Stream} stores a single read-only forensic data
set. For example, this stream might contains a hard disk image, a
memory image or a network capture (in PCAP format). Image streams have
an \texttt{aff2:type} attribute of \texttt{image}.

Storage for the data is done by using multiple data segments stored on
various volumes. Data segment URNs are derived by appending an 8
digit, zero padded decimal integer representation of an incrementing
id to the stream URN (e.g. ``urn:aff2:83a3d6db-85d5/00000032''). Each
data segment is called a bevy and stores a number of compressed chunks
back to back. The chunk index segment is a segment containing a list
of relative offsets to the beginning of each chunks within the
bevy. The chunk index segment URN is derived  by appending the bevy
URN with ``.idx''. This is illustrated in Figure \ref{image_stream}.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{image_segments.eps}}
  \caption{The structure of Image Stream Bevies. Each bevy is a
collection of compressed chunks stored back to back. Relative chunk
offsets are stored in the chunk index segment.}
  \label{image_stream}
  \end{center}
\end{figure}

Image streams specify the \texttt{chunk\_size} attribute, as the
number of image bytes each chunk contains (typically chunk sizes are
32kb). Also specified is the \texttt{chunks\_per\_segment} attribute
which specifies how many chunks are stored in each bevy. Each chunk is
compressed individually using the zlib compress algorithm. This
general structure of storing chunks within larger segments similar to
the technique used by the Expert Witness file format used by
EnCase\cite{encase-3.0} and implemented by the open source
libewf\cite{libewf} package. The advantages of keeping smaller sized
chunks is the better match between requested size and the minimum size
required for decompression. This means that less data is needed to be
decompressed unnecessarily.

%% FIXME - hashing may apply to any segment not just bevies
%Bevies can optionally be individually hashed, allowing for detection
%of modifications. The attribute \texttt{hash\_type} specifies the type
%of hash used (implementations will typically support md5,sha1 and
%sha256, but can define others). The hashes are stored in any number of
%segments within the stream, of arbitrary name. The hash segment names
%is provided by using the \texttt{hash\_name} stream attribute.

%Hashes are calculated in an identical way to that defined by AFF1's
%crypto layer\cite{garfinkel:affcrypto}---that is, the full name of the
%segment (including stream component) is followed by a single NULL
%byte, followed by the (uncompressed) content of the chunk. The
%inclusion of the segment name ensure that even identical chunks in
%content will have a different hash and prevents reordering attacks.

%Each hash segment is a list of a chunk id encoded in 32 bit
%little endian integer format followed by the hash (the length of which
%depends on the hash itself). This approach is illustrated in Figure
%\ref{hash_index}.

%Hash segments can contain any number of chunk hashes in them. It is
%recommended that a hash segment be created when a volume is closed to
%cover all the chunks within that volume. This allows for hashes to be
%verified even if some volumes are missing. The AFFLIB implementation
%may update a hash of a chunk previously set by an earlier hash
%segment---this would be required when a chunk is updated. The hash
%segment can then be signed to assure the integrity of the updated segment.

%\begin{figure}[tb]
%  \begin{center}
%  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
%  \caption{The structure of a hashing segment. The attributes {\em
%  hash} and {\em hash\_type} declare the hash segment. The segment
%  contains a list of integer encoded chunk id followed by an unencoded
%  hash for each chunk.}
%  \label{hash_index}
%  \end{center}
%\end{figure}

%This design differs from AFF1, which creates a separate segment for
%each hash. AFF4 uses headers that are more verbose than the AFF
%segment headers. Therefore, creating a new segment for each hash value
%would result in too much overhead: Packing the hashes into a few
%segments it is possible to reduce this overhead to a minimum.

\subsection{The Map Stream}
\label{map_stream}
Linear transformations of data are commonplace in forensic
analysis. For example, a file is often simply a collection of bytes
drawn from an image, while a TCP/IP stream is simply a collection of
payloads from selected network packets. 
Sometimes the same data may be
viewed in a number of ways---for example a Virtual Address Space is a
mapping of the Physical Address Space through a page table
transformation.  Zero Storage Carving \cite{Meijer2006} is a way of
specifying carved files in terms of a sequence of blocks taken from
the image; Cohen extended this concept to an arbitrary
mapping function\cite{Cohen2007} which can be used to describe
arbitrary mappings of carved files within a single image.

In this work we extend the mapping function concept to allow a single
map to draw data from arbitrary streams (called {\em targets}). This
transform is implemented via the {\em Map stream}.

The mapping function is described in a segment named by appending a
``/map'' to the stream URN. The segment data consists of a series of
lines, each containing a stream offset, followed by a comma, and a
target offset. Offsets are encoded using decimal notation, and are
followed by a target URN.

@@

For example, consider the following map:
\begin{lstlisting}
0,0,urn:aff2:83a3d6db-85d5
4096,10000,urn:aff2:f901be8e-d4b2
8192,5000,urn:aff2:83a3d6db-85d5
\end{lstlisting}

To read this stream we satisfy read requests of offsets between 0 and
4095 in the stream from offset 0 to offset 4095 in
\emph{urn:aff2:83a3d6db-85d5}. Requests for bytes between 4096 and 8191 are
fetched from \emph{urn:aff2:f901be8e-d4b2} at offset 10000. Finally
bytes after 8192 (until the specified size of the stream) are fetched
from offset 5000 in \emph{urn:aff2:83a3d6db-85d5}.

@@

Where $mod$ is the modulus function and $floor$ signifies integer
division. For example consider Listing~\ref{map}, which corresponds to a 3
disk RAID-5 array.

\begin{lstlisting}[
	float=tb, caption=A Map stream that corresponds to a 3 disk
	RAID-5 array. The targets are URNs for the respective
        disks. Note that map coordinates are given in multiples of 
	block size.,
	label=map, language=]
aff4:block_size=64k 
aff4:stream_period=6 
aff4:target_period=3

@@

%\Subsection{The Overlay Stream}
%The Overlay Stream Is A Variant Of The {\Em Image} Stream That
%Provides A Layer Of Indirection Between Requests Made For Specific
%Bytes (Or Chunks Or Bevys) And The Actual Location Where The Requested
%Information Is Stored. The Main Utility Of The Overlay Stream Is In
%Providing Transparent Access To Legacy Forensic Formats, Such As Ewf.

%To Access These Legacy File Formats We Only Need To Have An Index Of
%Each Chunk Offset, And Then Directly Use The Overlayed Files.  The {\Em
%Ewf2aff} Tool Provides Such An Overlay Feature For Ewf Volumes By
%Employing Libewf To Scan The Internal Data Structures.

%Overlay Stream Require That A \Texttt{Chunk\_Size} Attribute Be Specified. In
%Addition The {\Em Compression\_Type} Attribute May Be Specified. The
%Attribute Should Contain An Integer Corresponding To A Suitable Zip
%Compression Scheme (E.G. Deflate). If Not Specified It Is Assumed
%That No Compression Is Used.

%The ``Target'' Stream Attribute May Be Specified Multiple Times And
%Refer To External Files (See Section \Ref{Target_Reference}). There
%Can Be A Number Of Overlay Segments Named ``Stream Name/Overlay.00''
%Which Are All Merged To A Single Overlay. If More Than One Overlay Is
%Present, The Attribute {\Em Overlays} Must Be Specify How Many.

%The Overlay Segment Contains A Series Of Lines With Comma Delimited,
%Decimal Encoded, 64 Bit Integers Representing Chunk Number, Target
%Offset, Target Length And Target Number. Target Number Is An Index
%Into The Targets Defined In The Attributes (Count Starts With Zero).
%Chunks Are Then Fetched From The Specified Target And Decompressed If
%Necessary.

%It Is Possible To Refer To The Current File By Providing A Target Of
%``File://.''. This Can Be Useful For Modifying An Existing Ewf File To
%Be A Aff4 File. Since Zip Files Are Normally Read From The End Of The
%File, But Ewf Are Both Read From The Front Of The File, It May Be
%Possible To Append A Aff4 Central Directory To The End Of An Ewf File
%Without Interfering With The Overlayed File. In This Case The
%Converted File Can Still Be Used As An Ewf File Without Change, But
%Could Also Be Used As An Aff4 File.

%In Image Streams Some Chunks May Compress Very Well. In That Case, The
%Overheads Introduced By The Zip File Header For Each Chunk In The
%Image Stream May Become Unacceptably High. In That Case Its Possible
%To Use An {\Em Overlay Stream} To Coalesce Chunks Into Larger
%Segments. This Approach Is Illustrated In Figure \Ref{Overlay}.

%\Begin{Figure}[Tb]
%  \Begin{Center}
%  \Mbox{\Epsfxsize=0.6\Columnwidth \Epsffile{Overlay.Eps}}
%  \Caption{The Use Of An Overlay Stream To Overlay An Image With
%Coalesced Chunks.}
%  \Label{Overlay}
%  \End{Center}
%\End{Figure}

%In The Above Figure A Segment Is Written To The Aff4 Volume Which
%Contains A Number Of Chunks Back To Back. An Overlay Stream Then Uses
%Direct References To Each Compressed Chunk Within The Coalesced
%Segment. The Target Is Then Specified As The Segment Name. Although
%This Technique Is Valid, It Is Generally Discouraged Since It Breaks
%The Appealing Simplicity Of The Standard Image Stream.  It Is No
%Longer Possible To Simply Unzip All The Chunks And Concatenate Them
%Together To Recover The Original Image.

\subsection{Encrypted Streams}
\label{crypted_stream}
Encryption is an important property in an evidence file format. In
particular, multiple streams may be present in the file set, and often
different access levels are desired. For example, for evidence set
containing both network captures and disk images it may be desirable
to limit access to streams based on legal authorizations, even though
the same set is distributed to a number of people.

@@

AFF4 therefore introduces a new encryption scheme, the Encrypted
Stream.  The Encrypted Stream provides transparent encryption and
decryption onto a single target stream. The target stream actually
stores the encrypted data, and read requests from the stream are
satisfied by decrypting the relevant data from this backing
stream.

The Encrypted Stream may contain any data at all. It is useful
however, to store an actual AFF4 volume within the Encrypted
stream. This provides block level encryption for the contained AFF4
volume. It is recommended that Encrypted Streams containing an AFF4
volume set an attribute \texttt{content-type=application/x-aff-volume}
on the container stream, to allow implementations to automatically
recognize volumes. This approach is illustrated in
Figure~\ref{crypted_fif}.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.8\columnwidth \epsffile{crypted.eps}}
  \caption{Embedding an encrypted AFF4 volume within an Encrypted
Stream. The container volume contains an encrypted stream backed by an
image stream. Once the encrypted stream is opened, the volume stored
on its image stream is accessible. Now it is possible to see the
secret image stream stored within the volume.}
  \label{crypted_fif}
  \end{center}
\end{figure}

@@

\subsection{Encryption Schemes}
Our implementation provides a number of encryption schemes.

@@

When creating a new encryption stream, a master key is generated using
the first 128 bits of a SHA1 hash of the Pre-Shared-Key appended to a
64 bit random salt:
\begin{eqnarray}
Key_{Master} = \left | SHA1(salt + PSK) \right | _{128}
\end{eqnarray}

@@

The block is then encrypted or decrypted using AES in CBC mode with
this IV.

@@

\item Our implementation supports simultaneous access by multiple
  readers and writers. Our system allows concurrency by locking the
  file and interleaving write operations. Managing simultaneous
  writers is simplified through our update strategy which only writes
  new segments to the file, never modifying old segments.

\item AFF4 ``properties'' have specific semantics above and beyond
  normal segments. For example, AFF4 calls for all of the property
  segments associated with a file to be merged together. Furthermore,
  when updating the properties in a multi-volume Set, it is useful to
  write the same properties to each volume to ensure at least some of
the stream can be used even if some volumes are lost. Such
functionality is not present in standard ZIP64 implementations.
\end{itemize}
